\documentclass[../main.tex]{subfiles}

\graphicspath{{\subfix{../images/}}}

\begin{document}

    \section{Character Stats}

        As a component of a character, a character's stats define \emph{what} they are in the world; what they're capable of, and what they aren't.

    \subsection{Core Stats}

        The core stats of a character are generalized parts of them: how strong, fast, and smart they are. These stats are used for generalistic scenarios that involve less tact and more brute force to solve a problem, and are often used to determine the magnitude of a particular check's effectiveness if it succeeds.

        The core stats are divided into two groups of three, the first being \emph{mental stats}:

    \begin{itemize}
        \item \textbf{Will}: General knowledge or common sense, ability to learn, and ability to resist urges. 
        \item \textbf{Perception}: Ability to see, smell, hear, and detect the environment and subtleties of the world.
        \item \textbf{Conscious}: Ability to process and understand information and speed of cognition.
    \end{itemize}

        The second group is the {\em physical stats}:

    \begin{itemize}
        \item \textbf{Power}: Strength and physical capability.
        \item \textbf{Speed}: Agility and swiftness to commit actions.
        \item \textbf{Forte}: Physical health, conditioning, and ability to resist malicious infection.
    \end{itemize}

    \subsection{Stat Resources}

        Stat resources are a measure of something that a character has or something they're enduring; things like a curse, wealth, or the support of the people, that's been abstracted into a stat. This can be extremely useful to quantify something not necessarily quantifiable and then to allow something quantified to become part of the system in a way that allows it to be used in rolls.
        
        What stat resources a character has and how they function are things to discuss with the GM and other players. If you feel uncomfortable with this mechanic or would rather use exact numbers or roleplay to represent these things, then feel free to ignore stat resources altogether.

        Stat resources are divided into \emph{finite} and \emph{infinite} stat resources. Finite stat resources are things like wealth, the favor of the people, rations for a journey, and so on. These are things that are used up as you use them. Infinite resources are things like a curse or the blessing of your patron. Infinite stat resources are more complicated, and aren't used up when you use them; usually, it's much the opposite.

        Finite stat resources generally aren't rolled alone. They're used to add onto and modify rolls, but at a cost. When rolling a finite stat resource, \emph{any failures} reduce the Quantity of the resource by 1. Once the Quantity hits 0, the stat resource cannot be used until its Quantity is increased to above 1. Finite resource stats might be like this:

            \begin{displayquote}
                \emph{Agnar }
            \end{displayquote}

        Infinite stat resources are more like core stats, they typically level up like traditional stats and are used to solve problems all on their own. Infinite resource stats might be like this: 

            \begin{displayquote}
                \emph{Agnar }
            \end{displayquote}

    \subsection{Skills}
\end{document}